ODGI: understanding pangenome graphs

### 2022-03-15 manuscript/fig_extract_viz_draw_position_untangle.tex
italianization

        \textbf{(b-e)} \textit{odgi viz} visualizations of the C4 pangenome graph, where 8 paths are displayed: 2 reference genomes (chm13 and grch38 on the top) and 6 haplotypes of 3 diploid individuals.

@@

        The top two reference genomes and 2 haplotypes (HG01952\#2) go from left to right, while 5 haplotypes go in the opposite direction, as indicated by the black color on their left.

@@

        \textbf{(h)} Visualization of \textit{odgi untangle} output in the C4 pangenome graph: the plots show the copy number status of the sequences in the C4 region with respect to the grch38 reference sequence, making clear, for example, that in HG00438\#2, the \textit{C4A} gene is missing (white in the middle of the red). \vspace{-1em}

>>>>

        \textbf{(b-e)} \textit{odgi viz} visualizations of the C4 pangenome graph, where 8 paths are displayed: 2 reference genomes (\REVIEWED{CHM13} and \REVIEWED{GRCh38} on the top) and 6 haplotypes of 3 diploid individuals.

@@

        The top two reference genomes and one haplotypes (HG01952\#2) go from left to right, while 5 haplotypes go in the opposite direction, as indicated by the black color on their left.

@@

        \textbf{(h)} Visualization of \textit{odgi untangle} output in the C4 pangenome graph: the plots show the copy number status of the sequences in the C4 region with respect to the \REVIEWED{GRCh38} reference sequence, making clear, for example, that in HG00438\#2, the \textit{C4A} gene is missing (no black lines in the region annotated in red). \vspace{-1em}

### 2022-03-15 manuscript/fig_tips.tex
italianization

\begin{figure}[ht!]
	\begin{subfigure}{\linewidth}
		\caption{}
		\centering
		% include second image
		\includegraphics[width=1.0\linewidth, trim=0cm 2cm 0cm 2cm]{fig/tips/chr8_chm13_beta_defensin_locus_odgi_tips_w50000_karyoploteR}
		\label{fig:tips-karyo}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
	\caption{}
	\centering
	% include second image
	\includegraphics[width=1.0\linewidth, trim=-10cm 6cm -3.4cm 3cm]{fig/tips/chr8_pan_fa_c3d3224_7748b33.395c7f4_smooth_og_Y_og_p23_1_og_90paths_m}
	\label{fig:tips-viz}
\end{subfigure}
	%	\includegraphics[width=\linewidth]{fig/metrics/chr4.pan.HTTex1.gfa.multiqc_odgi_stats.png}
	\caption{Visualizing the contigs of a beta-defensin locus pangenome graph. (\textbf{a}) Breakpoint ranges of the contigs in the beta-defensin locus pangenome graph of 90 haplotypes relative to CHM13. The \textit{odgi tips} results are visualized as the density of BED ranges across the whole \textit{p23.1} cytoband. The high peak at the beginning of \textit{p23.1} indicates that 39 contigs of the graph broke at this specific position relative to CHM13. (\textbf{b}) \textit{odgi viz} color by path depth visualization of the 90 longest paths in the beta-defensin locus pangenome graph including CHM13 and GRCH38 on the top: using the Spectra color palette with 4 levels of path depths. White, grey, red, and yellow indicate a path depth of 0, 1, 2, and greater than or equal 3, respectively. Coloring by path depth we can see a very high path depth region on the right part of the  \textit{p23.1}. In this complex region, lot's of contigs break as can be seen in (\textbf{a}). There is no high path depth in the left part of the \textit{p23.1} where a huge number of contigs break at one specific reference position. This could be due to the fact that during graph construction, very small repeats are resolved instead of collapsed, and are therefore not highlighted in the visualization.}
	\label{fig:tips}
\end{figure}

>>>>

\begin{figure}[ht!]
	\begin{subfigure}{\linewidth}
		\caption{}
		\centering
		% include second image
		\includegraphics[width=1.0\linewidth, trim=0cm 2cm 0cm 2cm]{fig/tips/chr8_chm13_beta_defensin_locus_odgi_tips_w50000_karyoploteR}
		\label{fig:tips-karyo}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
	\caption{}
	\centering
	% include second image
	\includegraphics[width=1.0\linewidth, trim=-10cm 6cm -3.4cm 3cm]{fig/tips/chr8_pan_fa_c3d3224_7748b33.395c7f4_smooth_og_Y_og_p23_1_og_90paths_m}
	\label{fig:tips-viz}
\end{subfigure}
	%	\includegraphics[width=\linewidth]{fig/metrics/chr4.pan.HTTex1.gfa.multiqc_odgi_stats.png}
	\caption{Visualizing the contigs of a beta-defensin locus pangenome graph. (\textbf{a}) Breakpoint ranges of the contigs in the beta-defensin locus pangenome graph of 90 haplotypes relative to CHM13. The \textit{odgi tips} results are visualized as the density of BED ranges across the whole \textit{p23.1} cytoband. The high peak at the beginning of \textit{p23.1} indicates that 39 contigs of the graph broke at this specific position relative to CHM13. (\textbf{b}) \textit{odgi viz} color by path depth visualization of the 90 longest paths in the beta-defensin locus pangenome graph including CHM13 and GRCh38 on the top: using the Spectra color palette with 4 levels of path depths. White, grey, red, and yellow indicate a path depth of 0, 1, 2, and greater than or equal 3, respectively. Coloring by path depth we can see a very high path depth region on the right part of the  \textit{p23.1}. In this complex region, lot's of contigs break as can be seen in (\textbf{a}). There is no high path depth in the left part of the \textit{p23.1} where a huge number of contigs break at one specific reference position. This could be due to the fact that during graph construction, very small repeats are resolved instead of collapsed, and are therefore not highlighted in the visualization.}
	\label{fig:tips}
\end{figure}
